Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Clique Neighbourhood Constraints

### 2021-01-15 paper/paper.tex
Tweaks

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that simple distance filtering is valid for all
homomorphisms. However, we also show that \emph{none} of the other strategies are valid for finding
homomorphisms where there is no injectivity requirement. Finally, we introduce a new filtering
technique that is based upon finding a maximum clique in the neighbourhood of each pattern vertex, and
prove that this technique is valid even in the general case. Although this new filtering technique
involves solving a large number of \NP-complete problems, we demonstrate that it is effective in
practice, particularly for the non-injective problem where filtering allows a constraint programming
algorithm to solve instances over eight hundred times faster in aggregate, over a collection of over
fourteen thousand standard benchmark instances. This shows, for the first time, that modern
constraint programming techniques can be practical for less constrained graph homomorphism finding
problems; previous algorithmic approaches have focused instead upon worst-case computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

@@

\begin{proposition}\label{proposition:degreends}
    Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
    $v$ be any vertex. Then:
    \begin{enumerate}
        \item $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
        \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
            not pointwise greater than or equal to its own.
    \end{enumerate}
\end{proposition}
Since subgraph isomorphisms are in particular locally injective, these invariants also hold for
subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}; indeed, every recent constraint
programming approach for subgraph isomorphism finding make use of degrees and neighbourhood degree
sequences. The simplest way for a constraint programming algorithm to use these results is as unary
constraints, which are propagated when domains are initialised, before search begins. For each
domain representing a pattern vertex, any value representing target vertex whose degree or
neighbourhood degree sequence is too low may immediately be rejected.

@@

\begin{proposition}\label{proposition:distances}For the problem of finding a subgraph isomorphism
    $i$ from a graph $G$ to a graph $H$, the following constraints are implied for any pair of
    vertices $v, w \in \vertexset(G)$:
    \begin{enumerate}
        \item The distance between $v$ and $w$ is at least the distance between $i(v)$ and $i(w)$
            \cite{DBLP:conf/cp/AudemardLMGP14}.
        \item If there are at least $k$ simple
            paths of length exactly $\ell$ between $v$ and $w$, then there must be at least $k$
            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$
            \cite{DBLP:conf/cp/McCreeshP15}.
    \end{enumerate}
\end{proposition}

@@

\begin{proposition}\label{proposition:paths}For the problem of finding a locally injective graph
    homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
    \vertexset(G)$, if there are at least $k$ simple paths of length exactly two between $v$ and
    $w$, then there must be at least $k$ simple paths of length exactly two between $i(v)$ and
    $i(w)$.
\end{proposition}

\begin{proof}Let $\{ x_1, \ldots, x_n \}$ be the intermediate vertices on the paths of length two
between $v$ and $w$. Observe that each $x_j$ is in the neighbourhood of $v$, and so must be mapped
    to different vertices. Thus each sequence $(i(v), i(x_i), i(w))$ gives a distinct simple path of
length two between $i(v)$ and $i(w)$ as required.\end{proof}

>>>>

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that simple distance filtering is valid for all homomorphisms.
However, we also show that \emph{none} of the other strategies are valid for finding homomorphisms
where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
based upon finding a maximum clique in the neighbourhood of each pattern vertex, and prove that this
technique is valid even in the general case. Although this new filtering technique involves solving
many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
in practice, particularly for the non-injective problem where filtering allows a constraint
programming algorithm to solve instances over eight hundred times faster in aggregate, over a
collection of over fourteen thousand standard benchmark instances. This shows, for the first time,
that modern constraint programming techniques can be practical for less constrained graph
homomorphism finding problems; previous algorithmic approaches have focused instead upon worst-case
computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

@@

\begin{proposition}[neighbourhood degree sequences are preserved]\label{proposition:degreends}
    Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
    $v$ be any vertex. Then:
    \begin{enumerate}
        \item $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
        \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
            not pointwise greater than or equal to its own.
    \end{enumerate}
\end{proposition}
Since subgraph isomorphisms are in particular locally injective, these invariants also hold for
subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}; indeed, every recent constraint
programming approach for subgraph isomorphism finding make use of degrees and neighbourhood degree
sequences. The simplest way for a constraint programming algorithm to use these results is as unary
constraints, which are propagated when domains are initialised, before search begins. For each
domain representing a pattern vertex, any value representing a target vertex whose degree or
neighbourhood degree sequence is too low may immediately be rejected.

@@

\begin{proposition}[paths are preserved by subgraph isomorphisms]\label{proposition:distances}For
    the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph $H$, the following
    constraints are implied for any pair of vertices $v, w \in \vertexset(G)$:
    \begin{enumerate}
        \item The distance between $v$ and $w$ is at least the distance between $i(v)$ and $i(w)$
            \cite{DBLP:conf/cp/AudemardLMGP14}.
        \item If there are at least $k$ simple
            paths of length exactly $\ell$ between $v$ and $w$, then there must be at least $k$
            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$
            \cite{DBLP:conf/cp/McCreeshP15}.
    \end{enumerate}
\end{proposition}

@@

\begin{proposition}[paths of length two are preserved by locally injective
    homomorphisms]\label{proposition:paths}For the problem of finding a locally injective graph
    homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
    \vertexset(G)$, if there are at least $k$ simple paths of length exactly two between $v$ and
    $w$, then there must be at least $k$ simple paths of length exactly two between $i(v)$ and
    $i(w)$.
\end{proposition}

\begin{proof}Let $\{ x_1, \ldots, x_n \}$ be the intermediate vertices on the paths of length two
between $v$ and $w$. Observe that each $x_j$ is in the neighbourhood of $v$, and so must be mapped
    to different vertices. Thus each sequence $(i(v), i(x_j), i(w))$ gives a distinct simple path of
length two between $i(v)$ and $i(w)$ as required.\end{proof}
